---
title: Melbourne Housing
author: Fabianne Monserrat
output: github_document
---

# Assignment Project - STA 4102

## Wine Data

#### Open and read .csv file

```{r}
Data = read.csv("melb_data.csv", header = TRUE)
```

```{r}
dim(Data)
sum(is.na(Data))
Data = na.omit(Data)
sum(is.na(Data))
```

```{r}
library(dplyr)
glimpse(Data)
```

```{r}
sapply(Data, class)
```

Feature explanation:

#### Exploratory Analysis / Descriptive Statistics

```{r}
summary(Data$Price) 
```

```{r}
pairs(select_if(Data, is.numeric))
```

```{r}
library(corrplot)
corrplot(cor(select_if(Data, is.numeric)))
```

```{r}
cor(select_if(Data, is.numeric))
```

```{r}
hist(Data$Price)
```

#### Training

```{r}
library(caret)
set.seed(123)
trainIdx = createDataPartition(Data$Price, p = 0.7, list = FALSE, times = 1)
TrainingData = Data[ trainIdx ,]
TestingData = Data[ -trainIdx ,]
dim(TrainingData)
dim(TestingData)
```

### Linear Regression

```{r}
model_full = lm(Price ~ Rooms + Bathroom + Type + Distance + Car + Landsize + BuildingArea + YearBuilt, data = TrainingData)
summary(model_full)
```

Model R\^2 = 0.5557, 55.57% of variability in data explained by this model.

```{r}
PredictionFull = predict(model_full, newdata = TestingData)
MSE_Full = mean((PredictionFull - TestingData$Price)^2)
RMSE_Full = sqrt(MSE_Full)
RMSE_Full
```

```{r}
sqrt(mean(model_full$residuals^2))
```

```{r}
# Model without Type, Car, Landsize
model = lm(Price ~ Rooms + Bathroom + YearBuilt + Distance + BuildingArea, data = TrainingData)
summary(model)
```

```{r}
Prediction = predict(model, newdata = TestingData)
MSE = mean((Prediction - TestingData$Price)^2)
RMSE = sqrt(MSE)
RMSE
```

```{r}
sqrt(mean(model$residuals^2))
```

RMSE =

Original model RMSE =

Considering "HeartDisease" takes on either the value of 0 or not. The model RMSE for training and testing is not too far off, with the original having RMSE = 5977.949 (a difference of 218.328) therefore the model is pretty good. (Model is not overfit bc RMSE's are similar, and does perform consistently with new data)

### Decision Tree

```{r}
library(tree)
Tree = tree(Price ~ Rooms + Bathroom + YearBuilt + Distance + BuildingArea, data = TrainingData)
plot(Tree)
text(Tree, cex = .75)
summary(Tree)
```

```{r}
print("Making predictions for the following 5 houses:")
print(head(TrainingData))

print("The predictions are")
print(predict(Tree, head(TrainingData)))

print("Actual Price")
print(head(TrainingData$Price))
```

```{r}
PredictionTree = predict(Tree, newdata = TrainingData)
MSE = mean((PredictionTree - TestingData$Price)^2)
RMSE = sqrt(MSE)
RMSE
```

```{r}
library(rpart)
library(rpart.plot)
tree2 = rpart(Price ~ Rooms + Bathroom + YearBuilt + Distance + BuildingArea, data = TrainingData, control = rpart.control(minsplit = 2, cp = 0.005, maxdepth = 30))
# summary(tree2)
best2 = tree2$cptable[which.min(tree2$cptable[,'xerror']),'CP']
prunetree2 = prune(tree2, cp = best2)
prp(prunetree2)
```

```{r}
print("Making predictions for the last 5 houses:")
print(head(TrainingData))

print("Predicted Price:")
print(predict(tree2, tail(TrainingData)))

print("Actual Price:")
print(tail(TrainingData$Price))
```

```{r}
PredictionTree2 = predict(tree2, newdata = TrainingData)
MSE = mean((PredictionTree2 - TestingData$Price)^2)
RMSE = sqrt(MSE)
RMSE
```

RMSE = 0.4835247 (tree 1)

RMSE = 0.4938585 (tree 2)

The weighted average error between predictions and actual values is 0.940, which is only decent considering the range of quality in the data set is from 3 - 8.

### Random Forest

```{r}
library(randomForest)
rf1 = train(Price ~ Rooms + Bathroom + YearBuilt + Distance + BuildingArea, method = "rf", data = Data, trControl = trainControl(method = 'cv', number = 3))
summary(rf1)
```

```{r}
library(ROCR)
prediction1 = predict(rf1, newdata = TestingData)
```

```{r}
rf = randomForest(Price ~ Rooms + Bathroom + YearBuilt + Distance + BuildingArea, data = Data, importance = TRUE, mtry = 2)
vi = varImpPlot(rf)
```

```{r}
rf
```

```{r}
summary(rf)
```
